---
title: "SDS 192 Mini-Project 2"
subtitle: "SDS 192"
author: 
- name: Kat Brady
  affiliation: Smith College
- name: Ellie Zdancewic
  affiliation: Smith College
date: "`r format(Sys.Date(), '%B %e, %Y')`"
editor: 
  mode: source
output: 
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: true
    theme: lumen
    df_print: paged
    code_folding: show
---

```{r setup}
library(tidyverse)
library(fec20)
```

```{r helper-chunk}
# Look at help file
?contributions

# Note that the contributions data frame only contains the first 1000 rows:
contributions

# Download full contributions data from web and overwrite contributions using:
# read_all_*() function.
# You will need to install the usethis package for this line of code to work
contributions <- read_all_contributions()

# Now contributions has full 887,823 rows
contributions
```

```{r analysis-1}
#Kat's work in below chunks---------
```

```{r}
#making the data set smaller and filtering for relevant information
condensed_contributions<-contributions|>
  select(transaction_tp, entity_tp, name, transaction_amt,cand_id)

condensed_candidates<-candidates|>
  select(cand_id, cand_name, cand_pty_affiliation, cand_office)|>
  filter(cand_office=="P", cand_pty_affiliation%in%c("REP", "DEM"))
#joining the data with some data from candidates to add the candidates' names
contributions2<-condensed_contributions|>
  inner_join(condensed_candidates,
             by=c("cand_id"))
```

```{r}
#filtering for just Trump and Biden
contributions2<- contributions2|>
  filter(cand_id%in%c("P80001571","P80000722"))|>
  select(transaction_tp,entity_tp,name,transaction_amt,cand_id,cand_name)
```

```{r}
#Recoding the names of transactions to be more reader-friendly
contributions2$transaction_tp<-
  recode(contributions2$transaction_tp,
         "24E"="In Favor",
         "24K"="In Favor",
         "24A"="Against",
         "24C"="In Favor",
         "24N"="Against",
         "24F"="In Favor",
         "24Z"="In Favor")
```

```{r}
#This combines the multiple contributions made by one group so that the total amount contributed by one group is shown rather than every individual donation being shown differently
groupcontributions<- contributions2|>
  group_by(transaction_tp, entity_tp, name, cand_name)|>
  summarize(total=sum(transaction_amt))
```

```{r}
#To see the total quantity contributed to and against Biden And Trump by organizations, this last summary move is completed
sumcontributions<- groupcontributions|>
  group_by(transaction_tp, cand_name)|>
  summarize(total=sum(total))
```

```{r}
sumcontributions|>
ggplot(aes(x=total, y=cand_name, fill=transaction_tp))+
  geom_col()+
  coord_flip()+
  labs(title="Total Contributed to Leading Candidates by Groups*",
       subtitle="*In this case, 'groups' refers to non-individuals, including
committees, candidates, candidate committees, PACs, individual 
contributions to committees, and organizations",
x="Total (In USD)",
y="Candidate",
fill="Transaction",
caption="Source:
https://www.fec.gov/campaign-finance-data/contributions-committees-candidates-file-description/")+
scale_fill_manual(values=c("#ab400c","#65995c"))
```

```{r analysis-2}
# Ellie's Work------------------------------------------------------------------------------------- 
```

```{r}
#checking if there are any IND values in our dataset 
contributions_condensed_ind <- condensed_contributions |>
  filter(entity_tp %in% c("IND")) 
glimpse(contributions_condensed_ind)

```

```{r}
#Joining contributions with cand_id to match up Candidate names with Candidate IDs 
contributions3<-contributions|>
  inner_join(condensed_candidates,
             by=c("cand_id"))
```

```{r}
#creating a dataframe of only IND values regardless of amount contributed
contributions3_ind <- contributions3 |>
  filter(entity_tp %in% c("IND")) 
glimpse(contributions_condensed_ind)
```

```{r}
#filtering for contributions to Trump and Biden
contributions3_ind <- contributions3_ind |>
  filter(cand_name %in% c("BIDEN, JOSEPH R JR", "TRUMP, DONALD J.")) 
```

```{r}
#selecting only for relevant columns 
contributions3_ind<- contributions3_ind|>
  select(transaction_tp, name, city, state, transaction_amt,cand_name)
```

```{r}
#Recoding the names of transactions to be more reader-friendly
contributions3_ind$transaction_tp<-
  recode(contributions3_ind$transaction_tp,
         "24E"="In Favor",
         "24K"="In Favor",
         "24A"="Against")
```

```{r}
#filtering for only contributions in favor of the candidate
contributions3_ind <- contributions3_ind |>
  filter(transaction_tp %in% c("In Favor")) 
```

```{r}
#doesn't work yet 
#replacing instances of double commas with a single comma 
contributions3_ind$name <- contributions3_ind$name |> 
  str_replace_all(pattern = c(',,' = ","))
```

```{r}
#combining donors of the same name into one observation 
contributions3_ind_grouped<- contributions3_ind|>
  group_by(name, cand_name) |>
  summarise(total_amt = sum(transaction_amt))
```

```{r}
#creating my data visualization 
contributions3_ind_grouped |> 
  ggplot(aes(x=total_amt, fill = cand_name)) +
    geom_bar() + 
    facet_wrap(vars(cand_name)) + 
    scale_x_binned(
      n.breaks = 3
    ) + 
    scale_fill_brewer(palette = "Set1", direction = -1) + 
  labs(
    x = "Dollar Amount of Contribution", 
    y = "Count",
    title = "Large vs. Small Individual Committee Donations",
    subtitle = "For Committees Benefiting the 2020 Election Campigns of Trump and Biden",
    fill = "Candidate Name",
    caption = "Source: https://www.fec.gov/campaign-finance-data/contributions-committees-candidates-file-description/,"
  )
    
```

```{r}
#counting the number of biden contributions vs trump contributions 
contributions3_ind_grouped |> 
  group_by(cand_name) |>
  summarise(
    n = n()
  )
#BIDEN, JOSEPH R JR   838
#TRUMP, DONALD J.     145
```

# Blog post

The Federal Election Commission (FEC) is tasked with campaign finance regulation in all United States elections. Since a political donation was considered protected speech in the landmark 2010 Supreme Court case, Citizens United v. FEC, it is important to carefully record donations to ensure that they accord with existing restrictions. The contributions dataset contains records of 887,823 donations from groups including candidates, committees, and individuals donating to committees. The amount, type of transaction, donor's name, city, state, zip code and recipient of the donation was also added. The question that we wanted to consider was: Based upon contributions from individuals to committees as well as any group (non-individual) to the two leading candidates in the 2020 election, is it possible to assess if one candidate was more widely disliked than the other, or if one candidate was more likely to receive certain amounts of financial support? Therefore, we decided to narrow down the data by focusing only on the 2020 Presidential General Election between Donald Trump and Joe Biden.

To first address the quantities and broad "types" of support (either in favor of or against) Trump and Biden, we created our first visualization. This data depicts all of the funds donated either to or against the Trump and Biden campaigns, having been cleaned and having had other candidates filtered out. The resulting visualization shows that Biden received more money in donations for his campaign than Trump did, appearing to be over double Trump's amount. Additionally, Trump had more money spent against his campaign than Biden did, although the quantities are more similar than the quantities spent in favor. Essentially, Biden fared better with the so-called "group" donations than Trump did. However, it is important to note that the dataset used in making the visualization does not include any donations from individuals directly to either campaign, so it does not tell the entire story of the 2020 campaign donations. Thus, we cannot generalize about the donations of individuals directly to campaigns or the broader pattern of donation beyond those provided by organizations, committees, and other similar groups.

Our second visualization interrogates the donation amounts of individuals donating to committees in support of Trump and Biden's 2020 campaigns. We divided donations into two bins, "small" donations (less than 10,000) and "large" donations. The resulting bar chart indicates that both candidates received far more small donations than large donations. This makes sense intuitively because it is unlikely that many Americans have the means to donate several thousand dollars to any one committee, regardless of how strongly they feel about the candidate. However, it appears that Biden received more large donations than Trump as well as far more donations from individuals donating to committees overall. This is potentially because our data did not include direct donations from individuals to candidates, which was included in the Individuals dataset. A possible explanation for the gap between Trump and Biden's total donations is that Trump received more direct donations, but more data analysis is required for this to be confirmed. 


Based upon both of our visualizations, it seems fair to conclude that Biden was better supported by organizations, committees, candidate committees, candidates, PACs, and individuals who donated to committees than Trump was. While we cannot make a decisive conclusion about whether one of the candidates was more likely to get larger vs smaller donations, we do see that Biden was more likely to get donations and had a higher total value of donations from the aforementioned groups. Additionally, since Trump got less money donated from individuals to committees favoring him and from groups as a whole, all in addition to having more money donated to oppose his election, it is reasonable to conclude that Trump was perhaps more widely disliked. Again, we cannot apply these generalizations to donations directly from individuals to the candidates, but the data supports our conclusion outside of that context. In general, Biden was more preferred than Trump within the previously established context. A possible ethical concern of joining data across data frames is that when using joins such as inner_join, there is a possibility that data will be omitted because it does not have a matching key in the other data frame. For example, when we joined the Candidates and Contributions data frames using Cand_id as our key, there is a possibility that there is a candidate without a candidate ID (perhaps a write-in candidate) that was not included in the joined table. When joining tables we must be careful that we don't accidentally exclude important data points.
